introduction

### 2019-03-15 introduction.Rmd
structured: motivation, problem, previous work

Australia has changed in many ways over the last two decades. Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics[^1]. At the same time, political power has moved back and forth between the two major parties. In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal). The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period. This study explores the question of what electoral characteristics relate to party preference, and whether their effects have changed over time.

Previous work on modelling voting outcomes as a function of aggregate socio-demographics have found that relationships do exist, and relatively good models can be fit. @Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998, @Stimson06, @Stimson09 and @Stimson12 conduct principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote. @Stimson09 model the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, which has a poor fit.

>>>>

Australia has changed in many ways over the last two decades. Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics[^1]. At the same time, political power has moved back and forth between the two major parties. In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal). The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period. This study explores the question of how electoral characteristics relate to party preference, and whether their effects have changed over time.

Joining aggregate socio-demographics information from the Census with federal election voting data is problematic, as there in an inherent asynchronicity in the two events - a Census occurs every five years, whereas an election usually occurs every three years. The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09). Furthermore, electoral boundaries change regularly, so spatial discrepencies also arise when matching electoral data. To obtain appropriate Census data for the 2004, 2007, 2010 and 2013 elections, we adopt impute electoral socio-demographics using spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation. The process of collection and wrangling is detailed in @CHAPTER2 and imputation process is outlined in @CHAPTER3.

Previous work on modelling voting outcomes with aggregate socio-demographics find that relatively good models can be fit. @Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998. @Stimson06, @Stimson09 and @Stimson12 conduct principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote. @Stimson09 also model the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, but had little success. Instead of analyzing a single election in isolation, this paper employs a consistent model framework across elections so that any temporal changes are observed.
